Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-10-31 paper/pool_model.tex
formatting first section

\subsection{Three-Pool Model}
We begin our discussion by considering a simplistic system where only one bacteria species is present $N=1$.

@@

This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\citep{rolfe_lag_2012} and poorly studied physiological and regulatory mechanisms~\citep{monod_growth_1949}.

@@

In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying process: adaptation, growth or death.
Motivated by this we divide a bacterial population into three pools (states)~(\ref{fig:SchematicRep}), defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.

@@

To determine the system of equations which describe the system's dynamics, we consider the flows between the different states.

@@

For now, we assume for that there is no back flow between growth state $G$ and lag-state $L$.

@@

    G &\stackrel{\mu'}{\longrightarrow} & D

@@

The control of the population density with respect to the available nutrition is implemented using cell communication processes called quorum sensing.

@@

We will not dive into the processes of quorum sensing and try to keep it as simple as possible.
We aim to include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:

@@

    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G.

@@

@@

Let say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g.,, mol).

@@

This can be simplified by exploiting the conservation of the resource pool:

@@

@@

The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g.,, occupying space or consuming important nutrients.
If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantages of having one parameter less.\\

@@

However, in the case of time dependent parameters, e.g.,, for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.

>>>>

\subsection{Three-pool model}
We begin our discussion by considering a simplistic system where only one species is present $N=1$.

@@

This so-called preparation before the exponential growth includes synthesis of the needed compounds, initiation of the transcription~\citep{rolfe_lag_2012} and other physiological and regulatory mechanisms~\citep{monod_growth_1949}.

@@

In the described simplified version of the real-life system three possible bacterial states are mentioned, corresponding to the underlying processes: adaptation, growth or death.
Motivated by this we divide a bacterial population into three pools (states) (Figure~\ref{fig:SchematicRep}), defining the fraction of the population in the lag phase $L(t)$, growing and dividing $G(t)$ and undergoing cell death $D(t)$.

@@

To determine the equations which describe the system's dynamics, we consider the flows between the different states.

@@

For now, we assume no back flow from the growth state $G$ to the lag-state $L$.

@@

    L &\stackrel{\lambda}{\longrightarrow} G\\
    L &\stackrel{\mu}{\longrightarrow} D\\
    G &\stackrel{\alpha}{\longrightarrow} 2G\\
    G &\stackrel{\mu'}{\longrightarrow} D

@@

The control of the population density with respect to the available nutrition by means of cell communication is called quorum sensing.

@@

Instead of modeling complex signalling networks, we choose a simpler approach to describe the dependence of growth on the nutrient availability.
We include a resource pool which is consumed by the growing bacteria using the second order kinetic reaction:

@@

    G + R  \stackrel{\alpha_0}{\longrightarrow} 2G

@@

allowing us to keep a minimal number of parameters.

@@

Let's say $G=N_B/V$ and $[G]=1/V$, where $N_B$ is the number of bacteria and $V$ is the volume.
Then $R=N_R/V$ and $[R]=1/V$, where $N_R$ is the number of resource molecules (or any appropriate unit, e.g., mol).

@@

This can be simplified by exploiting the conservation of the resource pool

@@

        \dot{D} &= \mu  L + \mu' G.

@@

The relevant question here is whether cells in the pool $D$ still contribute to the limitation of the resources by, e.g., occupying space or consuming important nutrients.
If yes, the three pool model would be advantageous, if not the two pool model is sufficient with the advantage of having one parameter less.\\

@@

However, in the case of time dependent parameters, e.g., for a dependence on a non-static, dynamic temperature, it appears to be not advantageous to use the analytical results for $L$ and $D$ and work instead with all three equations for $L$, $G$, and $D$.
